Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-09 paper/main.tex
Updates

% Approaches to causal discovery
% have been developed within both the Directed Acyclic Graphs (DAGs) and
% Structural Equation Models (SEMs) frameworks, each adapting a different
% approach.

@@

\begin{enumerate}
%     \item We propose a novel measure of conditional association for mixed data
% 	    based on canonical correlations
% 	    (Section~\ref{sec:mixed_association}). This measure generalizes
% 	    several commonly used special case measures of association to mixed
% 	    data.
    \item We develop an iterative method based on CI testing to assist researchers in 
	    constructing DAGs using their domain knowledge (Section~\ref{sec:modification}).
    \item We develop a ranking method to prioritize potential modifications in the DAG that 
    	    can help in improving the model's fit quickly (Section~\ref{sec:ranking}).
    \item We provide a web-based interactive tool and a Python package to allow 
	    researchers to easily apply this method to their datasets (Section~\ref{sec:web}).
    \item Lastly, we compare the manual DAG construction method with automated
	    causal discovery algorithms (Section~\ref{sec:empirical}).
\end{enumerate}

% The rest of the paper is structured as follows. In
% Section~\ref{sec:background}, we give a background on the commonly used
% measures of association for various data types. In
% Section~\ref{sec:mixed_association}, we present our generalized measure of
% conditional association for mixed data. Section~\ref{sec:modification} provides
% details on the procedure for using the measure of association for constructing
% DAGs. Section~\ref{sec:web} presents our web-browser based tool and lastly in
% Section~\ref{sec:empirical}, we show some empirical results to compare this
% manual DAG construction approach to automated algorithms.

@@

\subsection{Measures of Conditional Association}
Our approach for ranking CI tests is based on a measure of conditional
association -- also known as partial association -- between variables.
Specifically, we are interested in quantifying the association between
variables $ X $ and $ Y $ when conditioned on a set of variables $ \bm{Z} $
(which may be empty, i.e., $ \bm{Z} = \emptyset $). Various measures of
conditional association have been used for this purpose depending on the type
of $ X $, $ Y $, and $ \bm{Z} $. These measures are often based on the effect
size of the CI test $ X \ci Y \rvert \bm{Z} $. In this section, we give an
overview of some of the commonly used measures for different data types.

>>>>

@@

\begin{enumerate}
    \item We develop an iterative method based on CI testing to assist researchers in 
	    constructing DAGs using their domain knowledge (Section~\ref{sec:modification}).
    \item We develop a ranking method to prioritize potential modifications in the DAG that 
    	    can help in improving the model's fit quickly (Section~\ref{sec:ranking}).
    \item We provide a web-based interactive tool and a Python package to allow 
	    researchers to easily apply this method to their datasets (Section~\ref{sec:web}).
    \item Lastly, we compare the manual DAG construction method with automated
	    causal discovery algorithms (Section~\ref{sec:empirical}).
\end{enumerate}

@@

\subsection{Measures of Conditional Association}
Our DAG structure learning approach relies on using CI tests and a measure of
conditional association between variables -- also known as partial association
-- between variables. Specifically, firstly we are interested in testing a CI
test: $ X \ci Y \bm{Z} $, i.e., whether the variables $ X $ and $ Y $ are
independent when conditioned on $ \bm{Z} $. Secondly, we are interested in
quantifying the strength of relationship between $ X $ and $ Y $ conditioned on
$ \bm{Z} $. We can quantify this using the effect size measures of the CI
tests. In this section, we give an overview of some of the commonly used CI
tests for various data types and their effect size measures.
